*Prochlorococcus marinus* responses to light and oxygen

### 2024-03-01 Code/Manuscript_O2.Rmd
doug scan of Manuscript_O2.Rmd edits, notes on edits missing

*Prochlorococcus marinus* comprises many strains, organized into clades, defined by 16S-23S intergenic transcribed ribosomal sequence signatures [@rocapGenomeDivergenceTwo2003]. The clades inhabit distinct ecological niches [@moorePhysiologyMolecularPhylogeny1998], originally defined by High-Light (HL) or Low-Light (LL).  Of at least 12 known *Prochlorococcus* genetic clades only 5 to date have cultured representatives; HLI, HLII, LLI, LLII/III and LIV [@billerProchlorococcusStructureFunction2015].

Current niches of *P. marinus* strains span ocean water columns [@veldhuisVerticalDistributionPigment1990; @chisholmNovelFreelivingProchlorophyte1988; @partenskyProchlorococcusMarinePhotosynthetic1999] and extend into regions with low dissolved oxygen concentrations [@partenskyProchlorococcusAdvantagesLimits2010; @goerickeNovelNicheProchlorococcus2000; @lavinNovelLineagesProchlorococcus2010; @ulloaCyanobacteriumProchlorococcusHas2021; @garcia-robledoCrypticOxygenCycling2017].

Low-Light clades thrive in deeper ocean waters, extending beyond 200 m in depth [@partenskyProchlorococcusMarinePhotosynthetic1999], where only ~1% of the surface irradiance penetrates, primarily in the blue (450 nm) to green (520 nm) spectral range [@holtropVibrationalModesWater2021]. Clade LLI includes cultured strain NATL2A, which prefers moderate irradiance typically between 30 and 100 m depth. Clades LLII and LLIII, including cultured strain SS120, are grouped together as the second oldest phylogenetic lineage diversifying in the *P. marinus* radiation, with an affinity for low light. Clade LLIV, including cultured strain MIT9313, falls near the base of the *Prochlorococcus* radiation, and is characterized by preference for low light, typical of depths from 120 m to 200 m [@partenskyProchlorococcusMarinePhotosynthetic1999]. LLIV members are, as yet, the only cultured strains to have been found in Oxygen Minimum Zones (OMZ).  Some, as yet, uncultured *P. marinus* strains in clades LLV and LLVI also thrive in OMZ of the subtropical Atlantic and Pacific Oceans, where dissolved oxygen concentrations [O~2~] can be less than 20 µM [@ulloaCyanobacteriumProchlorococcusHas2021; @lavinNovelLineagesProchlorococcus2010; @goerickeNovelNicheProchlorococcus2000; @garcia-robledoCrypticOxygenCycling2017; @johnsonEnergeticsGrowthKinetics1999]. *Prochlorococcus marinus* LL ecotypes may dominate the phytoplankton within OMZ [@lavinNovelLineagesProchlorococcus2010; @ulloaCyanobacteriumProchlorococcusHas2021; @partenskyProchlorococcusAdvantagesLimits2010], where they may be net O~2~ consumers [@partenskyComparisonPhotosyntheticPerformances2018].

HL clades are more evolved lineages, with reduced genome sizes in comparison to LL clades. High-Light clades are typically dominant picophytoplankters in near-surface, oligotrophic waters, characterized by high light levels. Clade HLI, represented by cultured strain MED4, is adapted to higher iron, and lower temperatures, and originated from 5 m depth in the Mediterranean Sea [@partenskyProchlorococcusMarinePhotosynthetic1999].  Clade HLII, adapted to higher iron, and higher temperatures, is the most abundant *P. marinus* clade in the North Atlantic and North Pacific Oceans, often constituting over 90% of the total population [@partenskyProchlorococcusMarinePhotosynthetic1999],  and are most numerous around 50 m depth [@partenskyProchlorococcusMarinePhotosynthetic1999]. Clade HLIII/IV is adapted to lower iron [@kentParallelPhylogeographyProchlorococcus2019; @johnsonNichePartitioningProchlorococcus2006; @zinserInfluenceLightTemperature2007].

@@

Our changing climate is rapidly altering conditions for these specialized clades of marine picophytoplankton. Predictions indicate a net global increase of *P. marinus* cell abundances of 29%, along with  poleward latitudinal shifts of at least 10° in marine phytoplankton niches by the end of this century [@flombaumPresentFutureGlobal2013; @bartonAnthropogenicClimateChange2016] in response to warming waters, with increases in *P. marinus* of approximately 50% in the more poleward regions of their distributions.

Near the equator, photoperiod remains nearly constant at the ocean surface, approximately 12 hours (h) of daylight and 12 h of darkness throughout the year. The effective length of the photoperiod does, however attenuate with depth as dawn and dusk light at depth drops below levels needed for biological processes. As *P. marinus* potentially expands its temperature-permissive niches poleward into temperate regions  [@flombaumPresentFutureGlobal2013; @bartonAnthropogenicClimateChange2016], it will encounter more
pronounced seasonal variations in photoperiod regimes both at surface and at depth, with potentially complex effects upon viable growth niches [@liDiatomGrowthResponses2017; @prezelinDielPeriodicityPhytoplankton1992]. For example, Vaulot *et al*. @vaulotGrowthProchlorococcusPhotosynthetic1995 showed that *Prochlorococcus* replication of DNA occurs in the afternoon, while cell division occurs at night.  To our knowledge, no study has as yet addressed *P. marinus* growth responses in relation to a  range of photoperiods.

Climate change is also rapidly changing ocean chemistry.  By the end of this century, surface ocean pH is projected to decline by 0.1 to 0.4  due to projected increases in carbon dioxide concentrations [@garcia-sotoOverviewOceanClimate2021]. Moreover, substantial changes in the global water cycle, leading to extensive changes in worldwide precipitation patterns, are affecting ocean salinity levels on a global scale, and ice melts due to rising temperatures are impacting salinity levels in the Arctic and Northwest Atlantic oceans [@leeRecognizingSalinityThreats2022]. Increasing sea temperatures are also causing decreases in [O~2~] across global oceans [@matearLongtermChangesDissolved2003], particularly toward the poles [@helmObservedDecreasesOxygen2011]. Warmer ocean waters increase stratification, and decrease oxygen solubility at the surface, which in turn decreases oxygen mixing downwards by ocean currents [@garcia-sotoOverviewOceanClimate2021]. Models predict that OMZ in the Pacific and Indian Oceans are expanding [@garcia-sotoOverviewOceanClimate2021; @buseckeDivergingFatesPacific2022]. In spite of these expanding OMZ, Busecke *et al*. [@buseckeDivergingFatesPacific2022] project that the cores of the OMZ, where the oxygen levels are lowest, may actually contract.

We used the OceansMap Protein Portal (OPP; https://proteinportal.whoi.edu/) [@saitoDevelopmentOceanProtein2021] to analyze the distribution of proteins from clades of *P. marinus* in samples taken across a range of [O~2~] and depth, which in turn correlates to peak light at the site of sampling. In parallel we analyzed the growth and physiological responses of representative strains from three clades of *P. marinus* under a matrix of [O~2~], light levels, spectral waveband ranges, and photoperiods, to approximate eco-physiological conditions representative of current and hypothetical future ocean zones. *Prochlorococcus marinus* MED4, a clade HLI strain, was isolated near the ocean surface (5m depth) of the Mediterranean Sea where [O~2~] is near saturation, light levels are high and spectral bias from full solar irradiance is minimal. *Prochlorococcus marinus* SS120, a clade LLIII strain, was isolated from the Sargasso Sea at a depth of 120 m, while *P. marinus* MIT9313, a clade LLIV strain, was isolated from the North Atlantic Gulf Stream at a depth of 135 m. At these depths, light attenuation and spectral shifts occur, resulting in low blue and green light, while [O~2~] varies from near-surface saturation levels to decreased concentrations, but does not necessarily decrease systematically with depth [@billheimerOxygenSeasonalityUtilization2021].

Genes encoding much of the photosynthetic apparatus is highly conserved in *Prochlorococcus* [@kettlerPatternsImplicationsGene2007], Photosynthetic organisms absorb light energy within the Photosynthetically Active Radiation (PAR) range, 350 to 700 nm, for photosynthesis [@morelAvailableUsableStored1978]. Photosynthetically Usable Radiation (PUR) represents the fraction of PAR that is absorbed by the pigments of a given photosynthetic organism [@morelAvailableUsableStored1978], taking into account the specific spectral wavebands these pigments absorb. *Prochlorococcus marinus* light-harvesting complexes show an absorption maxima of 442 nm for divinyl chlorophyll *a* and 478 nm for divinyl chlorophyll *b* [@goerickePigmentsProchlorococcusMarinus1992] allowing *P. marinus* to efficiently harvest blue light in the 400 nm to 500 nm range [@morelAvailableUsableStored1978] where only blue spectral wavelengths prevail in deep ocean habitats [@holtropVibrationalModesWater2021], thus photosynthetic efficiency may vary among clades [@hessPhotosyntheticApparatusProchlorococcus2001; @moorePhotophysiologyMarineCyanobacterium1999].  In *P. marinus* small cell diameters, from 0.5 to 0.7 µm [@chisholmProchlorococcusMarinusNov1992], and simple cell structures, minimize the complication of pigment package effect or intracellular self-shading [@morelProchlorococcusSynechococcusComparative1993] contributing to efficient optical absorption.

Given the different spectral light regimes typical of the niches of different ecotypes, expressing growth rates in terms of cumulative diel PUR might simplify different photoperiods, spectral bands, and PAR levels into a common parameter, making growth response comparisons across strains and different oxygen levels more accessible.  We also aimed to detect whether growth responses are driven simply by cumulative diel PUR, or whether specific photoperiods, spectral bands or PAR levels have independent, albeit interacting, effects on growth. We therefore analyzed growth rates in terms of cumulative diel PUR.

We discuss our findings in relation to analyses of genomic sequences [@changBRENDAELIXIRCore2021] across clades of *P. marinus*, showing that differences in the presence of genes encoding oxygen-dependent enzymes, and DNA repair enzymes, can help explain differential growth responses of strains under the matrix of light and [O~2~] conditions of this study.

@@

Metaproteomic datasets were obtained from the KM1128 entry in the BCO-DMO database (https://www.bco-dmo.org/deployment/59053) accessed via the OPP in June 2019. Datasets contained:
	i) Peptide sequences and sample identification (ID) number;
	ii) Sample ID number, station, depth in meters below the surface the sample was collected at, best-hit BLASTP protein and species annotation and the corresponding Uniprot Entry number for the identified proteins;
	iii) Sample station depth and [O~2~].  
	The depth and [O~2~] were joined to peptide sequence and BLASTP annotations by ID number, depth and station using *tidyverse* package [@R-tidyverse] running under R v4.1.3 and RStudio v2023.06.0 XXX Naaman to get location (https://posit.co/)  XXX. The resulting merged dataset was filtered for those *Prochlorococcus* peptides, detected from 0 to 300 m below the surface, annotated as a subunit of *Prochlorococcus* chlorophyll binding proteins (Pcb); Photosystem II (PSII); Cytochrome b~6~f (Cytb~6~f);  Photosystem I (PSI); NADPH Dehydrogenase (NDH); Plastoquinol Terminal Oxidase (PTOX); Plastocyanin (PC); Ferredoxin (Fd); Ribulose-1,5-bisphosphate oxygenase (RUBISCO);  ATP Synthase;  FtsH proteases (FtsH) or ribosomes. 
	Detected peptides were re-annotated for consistency and labelled, where feasible, according to strain, clade, subunit and protein complex. Full protein sequences corresponding to detected proteins were obtained from UniProt (https://www.uniprot.org/) and analyzed in Molecular Evolution and Genetic Analyses X (MEGAX) software  (https://www.megasoftware.net/). Sequences for proteins for each of the thirteen Prochlorococcus strains identified in the dataset were aligned with MUSCLE using UPGMA cluster method and a lambda of 24 with a -2.9 gap open penalty and 1.20 hydrophobicity multiplier.  Overall mean pairwise distance between protein sequences was determined using bootstrap variance estimation methods. Maximum likelihood phylogenetic trees were assembled using 1000 bootstrap replications with a 95% site coverage cut off. *Prochlorococcus* FtsH isoform identities, and functions, were inferred by sequence comparisons to the characterized four isoforms of FtsH protease of *Synechocystis* sp. PCC6803 (CITATONS).  Data for each strain was plotted against depth and  [O~2~] and sampling station.

@@

Controlled growth experiments were performed using MCMIX-OD or MC1000-OD PSI Multicultivators (Figure \@ref(fig:LabeledMC); PSI, Drásov, Czech Republic).  Each multicultivator individually controls 8 tubes at a common temperature of 22°C.  Each tube containing 70 mL of Pro99 media was inoculated with 10 mL of growing maintenance culture. In a factorial matrix design, each tube was then subject to an individual combination of sinusoidal photoperiod (4, 8, 12, 16  h);  reaching a peak PAR (30, 90, 180 µmol photons m^-2^ s^-1^), with defined spectral bandwidth (White LED, 660 nm, 450 nm).  [O~2~] levels (2.5 µM, 25 µM, 250 µM) were imposed by bubbling tubes with varying ratios of air and
Nitrogen (N~2~), with consistent 0.05% of Carbon Dioxide (CO~2~)
gas, delivered through a 0.2 μm sterile microfilter via a G400 gas mixing system (Qubit Systems Inc., Kingston, Ontario, Canada). [O~2~] in situ was verified using oxygen optodes (PyroScience,
Germany) inserted into tubes for
real-time measurements, with a temperature probe in
the aquarium of the bioreactor to correct [O~2~] measures for temperature
fluctuations. In addition, the Pyroscience software corrected [O~2~] based on the
salinity of the media (32 ppt). The flow rate of the gas mixture was controlled, but variations in bubbling speed, PAR and culture density affected the [O~2~] achieved in each tube. A low [O~2~] of 0.5 µM - 5 µM (reported as 2.5 µM hereafter),  was achieved by sparging with a gas mixture containing 99.95% N~2~ and 0.05%
CO~2~. An intermediate [O~2~] of 10 - 25 µM (reported hereafter as 25 µM) was achieved by sparging with a gas mixture containing 98.95% N~2~, 0.05% CO~2~ and 1% O~2~. A high O~2~ of 200 µM - 280 µM (reported hereafter as 250 µM) was achieved by sparging with lab air (78% N~2~, 21% O~2~, 1% Ar and 0.05% CO~2~).

@@

*In situ* measurements of Optical Density (OD) 680 nm, a proxy for cell suspension density, cell size dependent scatter and cell chlorophyll content;  and OD 720 nm, a proxy for cell suspension density and cell size dependent scatter, were recorded every 5 minutes over least 8 to 14 days,  depending on the
duration of the lag phase, if any. All data obtained from the Multicultivator were saved
as comma separated values files  (https://github.com/FundyPhytoPhys/prochlorococcus_o2/tree/main/Data/RawData).

@@

Figure \@ref(fig:deltaODLogGrowthPlot) is an example of chlorophyll specific
growth estimates fitted from the high resolution ΔOD measurements for
each tube in a Multicultivator. The residuals of the logistic growth
curve fit are shown and the growth spectral waveband is plotted and
illustrates the imposed PAR (µmol photons m^-2^ s^-1^) and photoperiod
(h). Annotated code for data import, transformations and analyses is available at https://github.com/FundyPhytoPhys/prochlorococcus_o2/tree/main/Data/RawData

A Generalized Additive Model (GAM) [@woodGeneralizedAdditiveModels2017] was applied to the
relation between chlorophyll-specific µ, d^-1^  to photoperiod and PAR level,  for each growth [O~2~] level, and for the blue and red wavebands for growth, for each *P.marinus* strain in this study. The R package *`mgcv`* [@R-mgcv] was used to model the growth rate with smoothing terms and indicate the 90, 50 and 10% quantiles for growth rate across the levels of factors. Only growth rate estimates for which the amplitude of standard error was smaller than 30% of the fitted growth rate were included in the GAM. Our priority was the effects of ecologically relevant blue light on growth trends.  We included analyses of growth responses to red light, which is not ecophysiologically relevant, but which might prove mechanistically informative [@murphyPhotoinactivationPhotosystemII2017].

@@

The applied photoperiods were delivered using the sine circadian light function of the PSI Multicultivator to simulate light exposure approximating sun rise through to sunset. The area under the sinusoidal curves is equivalent to the area of a triangular photoregime of equivalent photoperiod (Campbell, unpub), therefore the equation to determine the cumulative diel PUR (µmol photons m^-2^ d^-1^) is one half of the base (photoperiod) multiplied by the height (PUR) (Equation \@ref(eq:SumPURDay)); where PUR is the actinic absorbed light calculated from equation \@ref(eq:PUR) (µmol photons m^-2^ s^-1^), 3600 is the time conversion from seconds to hour and photoperiod is the imposed photoperiod (h).

@@

## Statistical analyses {.unnumbered}

@@

Using the dataset Omar *et al*. [@omarAnnotationGenesEncoding2023], we filtered the dataset for Enzyme Commission Numbers (EC numbers), or Kegg Orthology Numbers (KO numbers) identified by BRENDA  [@changBRENDAELIXIRCore2021] as 'natural substrates' for O~2~ in *P. marinus* (MED4, MIT9313, SS120, and NATL2A) in order to generate a subset containing only those orthologs encoding enzymes directly mediating O~2~ metabolism. We grouped orthologs together by EC number and their KO number and determined the occurrences of individual orthologs encoding each EC number, or KO number when EC number was not available, in a given strain. We merged the dataset with a list of enzyme Michaelis constant (K~m~) values from other organisms, as K~m~ values from *Prochlorococcus* were only available in the case of Ribulose bisphosphate carboxylase. XXX Possibly add the link to the full list of genes. Remove this once added XXXX

@@

```{r NATLProchloroPhotosynthDepthO2, fig.cap = NATLProchloroPhotosynthDepthO2_cap , fig.height = 20, fig.width = 10, echo = FALSE}

@@

Cumulative diel PUR can potentially collapse photoperiod, PAR and spectral wavebands to a common metric of photosynthetically active light absorbed per day. Cumulative diel PUR dose (µmol photons m^-2^ d^-1^) was calculated from the imposed PUR XXX should say imposed PAR?? or absorbed PUR ?? XXX (µmol photons m^-2^ s^-1^) and photoperiod (h). We plotted growth rates vs. cumulative diel PUR to determine whether growth is a simple response to diel PUR, across imposed photoperiods and spectral wavebands, or whether spectral wavebands or photoperiods have specific or interactive influences on growth beyond cumulative diel PUR.

>>>>

*Prochlorococcus marinus* comprises many strains, organized into clades, defined by 16S-23S intergenic transcribed ribosomal sequence signatures [@rocapGenomeDivergenceTwo2003]. The clades inhabit distinct ecological niches [@moorePhysiologyMolecularPhylogeny1998], originally defined as High-Light (HL) or Low-Light (LL).  Of at least 12 known *Prochlorococcus* genetic clades only 5 to date have cultured representatives; HLI, HLII, LLI, LLII/III and LIV [@billerProchlorococcusStructureFunction2015]. Current niches of *P. marinus* strains span ocean water columns [@veldhuisVerticalDistributionPigment1990; @chisholmNovelFreelivingProchlorophyte1988; @partenskyProchlorococcusMarinePhotosynthetic1999] and extend into regions with low dissolved oxygen concentrations [@partenskyProchlorococcusAdvantagesLimits2010; @goerickeNovelNicheProchlorococcus2000; @lavinNovelLineagesProchlorococcus2010; @ulloaCyanobacteriumProchlorococcusHas2021; @garcia-robledoCrypticOxygenCycling2017].

Low-Light clades thrive in deeper ocean waters, extending beyond 200 m in depth [@partenskyProchlorococcusMarinePhotosynthetic1999], where only ~1% of the surface irradiance penetrates, primarily in the blue (450 nm) to green (520 nm) spectral range [@holtropVibrationalModesWater2021]. Clade LLI includes cultured strain NATL2A, which prefers moderate irradiances typical of between 30 and 100 m depth. Clades LLII and LLIII, including cultured strain SS120, are grouped together as the second oldest phylogenetic lineage diversifying in the *P. marinus* radiation, with an affinity for low light. Clade LLIV, including cultured strain MIT9313, falls near the base of the *Prochlorococcus* radiation, and is characterized by preference for low light, typical of depths from 120 m to 200 m [@partenskyProchlorococcusMarinePhotosynthetic1999]. LLIV members are, as yet, the only cultured strains to have been found in Oxygen Minimum Zones (OMZ).  Some, as yet, uncultured *P. marinus* strains in clades LLV and LLVI also thrive in OMZ of the subtropical Atlantic and Pacific Oceans, where dissolved oxygen concentrations [O~2~] can be less than 20 µM [@ulloaCyanobacteriumProchlorococcusHas2021; @lavinNovelLineagesProchlorococcus2010; @goerickeNovelNicheProchlorococcus2000; @garcia-robledoCrypticOxygenCycling2017; @johnsonEnergeticsGrowthKinetics1999]. *P. marinus* LL ecotypes may indeed dominate the phytoplankton within OMZ [@lavinNovelLineagesProchlorococcus2010; @ulloaCyanobacteriumProchlorococcusHas2021; @partenskyProchlorococcusAdvantagesLimits2010], where they may be net O~2~ consumers [@partenskyComparisonPhotosyntheticPerformances2018].

HL clades are more recently branching lineages, with reduced genome sizes in comparison to LL clades. High-Light clades are typically dominant picophytoplankters in near-surface, oligotrophic waters, characterized by high light levels. Clade HLI, represented by cultured strain MED4, is adapted to higher iron, and lower temperatures, and originated from 5 m depth in the Mediterranean Sea [@partenskyProchlorococcusMarinePhotosynthetic1999].  Clade HLII, adapted to higher iron, and higher temperatures, is the most abundant *P. marinus* clade in the North Atlantic and North Pacific Oceans, often constituting over 90% of the total population [@partenskyProchlorococcusMarinePhotosynthetic1999],  and are most numerous around 50 m depth [@partenskyProchlorococcusMarinePhotosynthetic1999]. Clade HLIII/IV is adapted to lower iron [@kentParallelPhylogeographyProchlorococcus2019; @johnsonNichePartitioningProchlorococcus2006; @zinserInfluenceLightTemperature2007].

@@

Our changing climate is rapidly altering conditions for these specialized clades of marine picophytoplankton. Predictions indicate a net global increase of *P. marinus* cell abundances of 29% [@flombaumPresentFutureGlobal2013{}, along with  poleward latitudinal shifts of at least 10° in marine phytoplankton niches by the end of this century  [@bartonAnthropogenicClimateChange2016] in response to warming waters, with increases in *P. marinus* of approximately 50% in the more poleward regions of their distributions.

Near the equator, photoperiod remains nearly constant at the ocean surface, approximately 12 hours (h) of daylight and 12 h of darkness throughout the year. The effective length of the photoperiod does, however, attenuate with depth as dawn and dusk light at depth drops below levels needed for biological processes. As *P. marinus* potentially expands its temperature-permissive niches poleward into temperate regions  [@flombaumPresentFutureGlobal2013; @bartonAnthropogenicClimateChange2016], it will encounter more
pronounced seasonal variations in photoperiod regimes both at surface and at depth, with potentially complex effects upon viable growth niches [@liDiatomGrowthResponses2017; @prezelinDielPeriodicityPhytoplankton1992]. For example, Vaulot *et al*. @vaulotGrowthProchlorococcusPhotosynthetic1995 showed that *Prochlorococcus* replication of DNA occurs in the afternoon, while cell division occurs at night.  To our knowledge, no study has as yet addressed *P. marinus* growth responses in relation to a range of photoperiods.

Climate change is also rapidly changing ocean chemistry. By the end of this century, surface ocean pH is projected to decline by 0.1 to 0.4  due to projected increases in carbon dioxide concentrations [@garcia-sotoOverviewOceanClimate2021]. Moreover, substantial changes in the global water cycle, leading to extensive changes in worldwide precipitation patterns, are affecting ocean salinity levels on a global scale, and ice melts due to rising temperatures are impacting salinity levels in the Arctic and Northwest Atlantic oceans [@leeRecognizingSalinityThreats2022]. Increasing sea temperatures are also causing decreases in [O~2~] across global oceans [@matearLongtermChangesDissolved2003], particularly toward the poles [@helmObservedDecreasesOxygen2011]. Warmer ocean waters decrease oxygen solubility at the surface, and increase stratification, which in turn decreases oxygen mixing downwards by ocean currents [@garcia-sotoOverviewOceanClimate2021]. Models predict that OMZ in the Pacific and Indian Oceans are expanding [@garcia-sotoOverviewOceanClimate2021; @buseckeDivergingFatesPacific2022], although the cores of the OMZ, where the oxygen levels are lowest, may actually contract  [@buseckeDivergingFatesPacific2022].

We used the OceansMap Protein Portal (OPP; https://proteinportal.whoi.edu/) [@saitoDevelopmentOceanProtein2021] to analyze the distribution of proteins from clades of *P. marinus* in samples taken across a range of [O~2~] and depth, which in turn correlates to peak light at the site of sampling. In parallel we analyzed the growth and physiological responses of representative strains from three clades of *P. marinus* under a matrix of [O~2~], light levels, spectral waveband ranges, and photoperiods, to approximate eco-physiological conditions representative of current and hypothetical future ocean zones. *Prochlorococcus marinus* MED4, a clade HLI strain, was isolated near the ocean surface (5m depth) of the Mediterranean Sea where [O~2~] is near saturation, light levels are high and spectral bias from full solar irradiance is minimal. *Prochlorococcus marinus* SS120, a clade LLIII strain, was isolated from the Sargasso Sea at a depth of 120 m, while *P. marinus* MIT9313, a clade LLIV strain, was isolated from the North Atlantic Gulf Stream at a depth of 135 m. At these depths, light attenuation and spectral shifts occur, resulting in low blue light, while [O~2~] varies from near-surface saturation levels to decreased concentrations, but does not necessarily decrease systematically with depth [@billheimerOxygenSeasonalityUtilization2021].

Photosynthetic organisms absorb light energy within the Photosynthetically Active Radiation (PAR) range, 350 to 700 nm, for photosynthesis [@morelAvailableUsableStored1978]. Photosynthetically Usable Radiation (PUR) represents the fraction of PAR that is absorbed by the pigments of a given photosynthetic organism [@morelAvailableUsableStored1978], taking into account the specific spectral wavebands these pigments absorb. *Prochlorococcus marinus* Pcb light-harvesting complexes show an absorption maxima of 442 nm for divinyl chlorophyll *a* and 478 nm for divinyl chlorophyll *b* [@goerickePigmentsProchlorococcusMarinus1992] allowing *P. marinus* to efficiently harvest blue light in the 400 nm to 500 nm range [@morelAvailableUsableStored1978] corresponding to blue spectral wavelengths prevailing in deep ocean habitats [@holtropVibrationalModesWater2021].  In *P. marinus* small cell diameters, from 0.5 to 0.7 µm [@chisholmProchlorococcusMarinusNov1992], and simple cell structures, minimize the complication of pigment package effect or intracellular self-shading [@morelProchlorococcusSynechococcusComparative1993] contributing to efficient optical absorption, although photosynthetic efficiency may vary among clades [@hessPhotosyntheticApparatusProchlorococcus2001; @moorePhotophysiologyMarineCyanobacterium1999].

Given the different spectral light regimes typical of the niches of different ecotypes, expressing growth rates in terms of cumulative diel PUR might simplify different photoperiods, spectral bands, and PAR levels into a common parameter, making growth response comparisons across strains and different oxygen levels more accessible.  We  aimed to detect whether growth responses are driven simply by cumulative diel PUR, or whether specific photoperiods, spectral bands or PAR levels have independent, albeit interacting, effects on growth. We therefore analyzed growth rates in terms of cumulative diel PUR.

We discuss our findings in relation to analyses of genomic sequences [@changBRENDAELIXIRCore2021] across clades of *P. marinus*, showing that differences in the expression and presence of genes encoding protein turnover, oxygen-dependent enzymes, and DNA repair enzymes, can explain the differential growth responses of strains under the matrix of light and [O~2~] conditions of this study.

@@

Metaproteomic datasets were obtained from the KM1128 entry in the BCO-DMO database (https://www.bco-dmo.org/deployment/59053) accessed via the OPP in June 2019. Datasets contained:
	i) Peptide sequences and sample identification (ID) number;
	ii) Sample ID number, station, depth in meters below the surface the sample was collected at, best-hit BLASTP protein and species annotation and the corresponding Uniprot Entry number for the identified proteins;
	iii) Sample station depth and [O~2~].  
	The depth and [O~2~] were joined to peptide sequence and BLASTP annotations by ID number, depth and station using *tidyverse* package [@R-tidyverse] running under R v4.1.3 and RStudio v2023.06.0 XXX Naaman to get location (https://posit.co/)  XXX. The resulting merged dataset was filtered for those *Prochlorococcus* peptides, detected from 0 to 300 m below the surface, annotated as a subunit of *Prochlorococcus* chlorophyll binding proteins (Pcb); Photosystem II (PSII); Cytochrome b~6~f (Cytb~6~f);  Photosystem I (PSI); NADPH Dehydrogenase (NDH); Plastoquinol Terminal Oxidase (PTOX); Plastocyanin (PC); Ferredoxin (Fd); Ribulose-1,5-bisphosphate oxygenase (RUBISCO);  ATP Synthase;  FtsH proteases (FtsH) or ribosomes. 
	Detected peptides were re-annotated for consistency and labelled, where feasible, according to strain, clade, subunit and protein complex. Full protein sequences corresponding to detected proteins were obtained from UniProt (https://www.uniprot.org/) and analyzed in Molecular Evolution and Genetic Analyses X (MEGAX) software  (https://www.megasoftware.net/). Sequences for proteins for each of the thirteen *Prochlorococcus* strains identified in the dataset were aligned with MUSCLE using UPGMA cluster method and a lambda of 24 with a -2.9 gap open penalty and 1.20 hydrophobicity multiplier.  Overall mean pairwise distance between protein sequences was determined using bootstrap variance estimation methods. Maximum likelihood phylogenetic trees were assembled using 1000 bootstrap replications with a 95% site coverage cut off. *Prochlorococcus* FtsH isoform identities, and functions, were inferred by sequence comparisons to the characterized four isoforms of FtsH protease of *Synechocystis* sp. PCC6803 (CITATONS).  Data for each strain was plotted against depth and  [O~2~] and sampling station.

@@

Controlled growth experiments were performed using MCMIX-OD or MC1000-OD PSI Multicultivators (Figure \@ref(fig:LabeledMC); PSI, Drásov, Czech Republic).  Each multicultivator individually controls 8 tubes at a common temperature of 22°C.  Each tube containing 70 mL of Pro99 media was inoculated with 10 mL of growing maintenance culture. In a factorial matrix design, each tube was then subject to an individual combination of sinusoidal photoperiod (4, 8, 12, 16  h);  reaching a peak PAR (30, 90, 180 µmol photons m^-2^ s^-1^), with defined spectral bandwidth (White LED, 660 nm, 450 nm).  [O~2~] levels (2.5 µM, 25 µM, 250 µM) were imposed by bubbling tubes with varying ratios of air and
Nitrogen (N~2~), with consistent 0.05% of Carbon Dioxide (CO~2~) gas, delivered through a 0.2 μm sterile microfilter via a G400 gas mixing system (Qubit Systems Inc., Kingston, Ontario, Canada). [O~2~] in situ was verified using oxygen optodes (PyroScience, Germany) inserted into tubes for real-time measurements, with a temperature probe in
the bath of the bioreactor to correct [O~2~] measures for temperature fluctuations. In addition, the Pyroscience software corrected [O~2~] based on the salinity of the media (32 ppt). The flow rate of the gas mixture was controlled, but variations in bubbling speed, PAR and culture density affected the [O~2~] achieved in each tube. A low [O~2~] of 0.5 µM - 5 µM (reported as 2.5 µM hereafter),  was achieved by sparging with a gas mixture containing 99.95% N~2~ and 0.05% CO~2~. An intermediate [O~2~] of 10 - 25 µM (reported hereafter as 25 µM) was achieved by sparging with a gas mixture containing 98.95% N~2~, 0.05% CO~2~ and 1% O~2~. A high O~2~ of 200 µM - 280 µM (reported hereafter as 250 µM) was achieved by sparging with lab air (78% N~2~, 21% O~2~, 1% Ar and 0.05% CO~2~).

@@

*In situ* measurements of Optical Density (OD) 680 nm, a proxy for cell suspension density, cell size dependent scatter and cell chlorophyll content;  and OD 720 nm, a proxy for cell suspension density and cell size dependent scatter, were recorded every 5 minutes over least 8 to 14 days,  depending on the duration of the lag phase, if any. All data obtained from the Multicultivator were saved as comma separated values files  (https://github.com/FundyPhytoPhys/prochlorococcus_o2/).

@@

Figure \@ref(fig:deltaODLogGrowthPlot) is an example of chlorophyll specific growth estimates fitted from the high resolution ΔOD measurements for each tube in a Multicultivator. The residuals of the logistic growth curve fit are shown and the growth spectral waveband is plotted and illustrates the imposed PAR (µmol photons m^-2^ s^-1^) and photoperiod
(h). Annotated code for data import, transformations and analyses is available at https://github.com/FundyPhytoPhys/prochlorococcus_o2/  MAKE SURE RePOSTIORY IS PUBLICXXXX

A Generalized Additive Model (GAM) [@woodGeneralizedAdditiveModels2017] was applied to the relation of chlorophyll-specific µ, d^-1^  to photoperiod and PAR level,  for each growth [O~2~] level, and for the blue and red wavebands for growth, for each *P.marinus* strain in this study. The R package *`mgcv`* [@R-mgcv] was used to model the growth rate with smoothing terms and indicate the 90, 50 and 10% quantiles for growth rate across the levels of factors. Only growth rate estimates for which the amplitude of standard error was smaller than 30% of the fitted growth rate were included in the GAM. Our priority was the effects of ecologically relevant blue light on growth trends.  We also included GAM analyses of growth responses to red light, which is not ecophysiologically relevant, but which might prove mechanistically informative [@murphyPhotoinactivationPhotosystemII2017].

@@

The applied photoperiods were delivered using the sinusoidal circadian light function of the PSI Multicultivator to simulate light exposure approximating sun rise through to sunset. The area under the sinusoidal curves is equivalent to the area of a triangular photoregime of equivalent photoperiod (Campbell, unpub), therefore the equation to determine the cumulative diel PUR (µmol photons m^-2^ d^-1^) is one half of the base (photoperiod) multiplied by the height (PUR) (Equation \@ref(eq:SumPURDay)); where PUR is the actinic absorbed light calculated from equation \@ref(eq:PUR) (µmol photons m^-2^ s^-1^), 3600 is the time conversion from seconds to hour and photoperiod is the imposed photoperiod (h).

@@

@@

 We filtered the dataset of Omar *et al*. [@omarAnnotationGenesEncoding2023], for Enzyme Commission Numbers (EC numbers), or Kegg Orthology Numbers (KO numbers) identified by BRENDA  [@changBRENDAELIXIRCore2021] as 'natural substrates' for O~2~ in *P. marinus* strains (MED4, MIT9313, SS120, and NATL2A). We grouped orthologs together by EC number and their KO number and determined the occurrences of individual orthologs encoding each EC number, or KO number when EC number was not available, in a given strain. We merged the dataset with a list of enzyme Michaelis constant (K~m~) values from other organisms, as K~m~ values from *Prochlorococcus* were only available in the case of Ribulose bisphosphate carboxylase. XXX Possibly add the link to the full list of genes. Remove this once added XXXX
 XXXXADD METHODS for DNA REpair and for light dependent EnzymesXXXX

@@

XXXRewrite caption if we change figure versionXXX
```{r NATLProchloroPhotosynthDepthO2, fig.cap = NATLProchloroPhotosynthDepthO2_cap , fig.height = 20, fig.width = 10, echo = FALSE}

@@

Cumulative diel PUR can potentially collapse photoperiod, PAR and spectral wavebands to a common metric of photosynthetically active light absorbed per day. Cumulative diel PUR dose (µmol photons m^-2^ d^-1^) was calculated from the imposed PUR (µmol photons m^-2^ s^-1^) and photoperiod (h). We plotted growth rates vs. cumulative diel PUR to determine whether growth is a simple response to diel PUR, across imposed photoperiods and spectral wavebands, or whether spectral wavebands or photoperiods have specific or interactive influences on growth beyond cumulative diel PUR.
